---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions,echo=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r preprocess,results='hide'}
unzip('activity.zip')
activity <- read.csv('activity.csv')
activity$date <- as.Date(activity$date)
```

## What is mean total number of steps taken per day?

```{r stepsperday,fig.height=4}
steps_per_day <- aggregate(steps ~ date, activity, sum)
hist(steps_per_day$steps, xlab = 'Steps', main = 'Steps per day')
steps_mean <- format(mean(steps_per_day$steps))
steps_median <- format(median(steps_per_day$steps))
```

The mean of the total number of steps taken per day is `r steps_mean` and the median is `r steps_median`.

## What is the average daily activity pattern?

```{r avgactivity,fig.height=4}
avg_activity <- aggregate(steps ~ interval, activity, mean)
plot(avg_activity, type = 'l', main = 'Daily activity pattern')
sorted_activity <- avg_activity[order(avg_activity$steps, decreasing = TRUE),]
highest_activity_interval <- sorted_activity[1, 'interval']
```

Interval #`r highest_activity_interval` contains the maximum number of steps.

## Imputing missing values

```{r missingvalues}
missing_activity_cond <- is.na(activity$steps)
missing_activity_cnt <- sum(missing_activity_cond)
```

The total number of missing values is `r missing_activity_cnt`. Now we'll use the mean for each 5-minute interval strategy to fill in missing values.

```{r fillvalues,fig.height=4}
filled_activity <- as.data.frame(activity)
for (i in 1:nrow(filled_activity)) {
    if (is.na(filled_activity[i, 'steps'])) {
        interval <- filled_activity[i, 'interval']
        filled_activity[i, 'steps'] <- avg_activity[avg_activity$interval == interval, 'steps']
    }
}
steps_per_day2 <- aggregate(steps ~ date, filled_activity, sum)
hist(steps_per_day2$steps, xlab = 'Steps', main = 'Steps per day (filled dataset)')
steps_mean <- format(mean(steps_per_day2$steps))
steps_median <- format(median(steps_per_day2$steps))
```

The mean of the total number of steps taken per day is `r steps_mean` and the median is `r steps_median`. The median differs from the first part of the assignment, and it've become equal to the mean value.

## Are there differences in activity patterns between weekdays and weekends?

```{r weekendpatterns,fig.height=4}
library(lattice)
filled_activity$day_type <- factor(sapply(filled_activity$date, function(x) {
    if (weekdays(x) %in% c('Saturday', 'Sunday')) {
        'weekend'
    } else {
        'weekday'
    }
}))
avg_activity_day_type <- aggregate(steps ~ interval + day_type, filled_activity, mean)
xyplot(steps ~ interval | day_type, avg_activity_day_type, type = 'l', layout = c(1,2))
```